\mode*

% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section{Book-Keeping}

\subsection{Double-Entry Book-Keeping}

\begin{frame}
  \begin{itemize}
    \item The banks are one of the oldest institutions with a need for strict 
      accountability.

    \item The main tools developed for this purpose is double-entry 
      book-keeping.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{definition}[Double-entry book-keeping]
    \begin{itemize}
      \item Add one entry of \(x\) and one of \(-x\).
      \item Invariant of zero (\(x+(-x) = 0\)).
    \end{itemize}
  \end{definition}

  \begin{example}
    \begin{itemize}
      \item All books should be balanced.

      \item A transfer from one account to another must be a credit in one 
        account and a debit in the other.

      \item I.e.\ when adding them up they equal zero.
    \end{itemize}
  \end{example}
\end{frame}

%\begin{frame}
%  \begin{remark}
%    \begin{itemize}
%      \item This principle of keeping a balance of constant zero can be 
%        tranferred to other principles.
%    \end{itemize}
%  \end{remark}
%
%  \begin{example}
%    \begin{itemize}
%      \item E.g.\ for each log-in there should be a log-out.
%
%      \item If the difference of number of log-ins \(L_i\) for a user and the 
%        number of log-outs \(L_o\) is zero (\(L_i-L_o = 0\)), then the user is 
%        not currently logged-in.
%
%      \item Hence, the user shouldn't be able to post a comment when the system 
%        is in this state.
%    \end{itemize}
%  \end{example}
%\end{frame}
%
%\begin{frame}
%  \begin{remark}
%    \begin{itemize}
%      \item Note that you shouldn't use the book-keeping system to keep track of 
%        whether a user is logged-in or not.
%
%      \item You can use more efficient mechanisms for that.
%
%      \item But the account should be kept for future reference, in case 
%        something bad happens, then you should be able to see what really 
%        happened.
%    \end{itemize}
%  \end{remark}
%\end{frame}

\subsection{Separation of Duties}

\begin{frame}
  \begin{definition}[Separation of duties]
    \begin{itemize}
      \item Two or more entities must collude to break the policy.
      \item Two classes: \emph{dual control} and \emph{functional separation}.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{example}[Dual control]
    \begin{itemize}
      \item Two or more staff members must act together to authorize 
        a transaction.
    \end{itemize}
  \end{example}

  \pause{}

  \begin{example}[Dual control on film]
    \begin{itemize}
      \item Two guys in a nuclear weapons silo.
      \item Two keys too far from each other for one to turn simultaneously.
      \item Both staffers must agree to turn the keys.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}[Functional separation]
    \begin{itemize}
      \item Two or more staff members must act on the transaction at different 
        points in the transaction path.
    \end{itemize}
  \end{example}
  
  \pause{}

  \begin{example}[Functional separation]
    \begin{itemize}
      \item Developer team writes the code.
      \item System administrators deploy it.
      \item Auditors verify it.
    \end{itemize}
  \end{example}
\end{frame}

\subsection{Clark-Wilson Security Policy Model}

\begin{frame}
  \begin{idea}[Clark-Wilson Security Policy Model]
    \begin{itemize}
      \item Is a model for securely implementing a security policy.

      \item Ensures \emph{internal consistency}, i.e.\ properties of the internal 
        state of the system.

      \item Allows for \emph{external consistency}, i.e.\ the relation of the 
        internal state of the system to the real world.

      \item The latter must be enforced by auditing.
    \end{itemize}
  \end{idea}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Mechanisms for enforcing integrity of the system are:
      \begin{itemize}
        \item Well-formed transactions

        \item Separation of duties
      \end{itemize}
  \end{itemize}

  \pause{}

  \begin{definition}[Well-formed transactions]
    \begin{itemize}
      \item A limited set of functions can manipulate an object.
      \item Users have access to these functions, not the objects.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{block}{Requirements}
    \begin{enumerate}
      \item Subjects must be authenticated.
      \item Objects can be manipulated only by a restricted set of functions.
      \item Subjects can execute only a restricted set of functions.
      \item A proper audit log must be maintained.
      \item The system has to be certified to work properly.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}
  \begin{definition}[Unconstrained data item, UDI]
    \begin{itemize}
      \item Input from outside the system.
      \item From outside the control of the system.
      \item It can be anything!
    \end{itemize}
  \end{definition}

  \pause{}

  \begin{definition}[Constrained data item, CDI]
    \begin{itemize}
      \item Objects (data) inside the system.
      \item This is under the system's control.
      \item This is well-formed.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item UDIs must be converted to CDIs.
      \item This is a critical part of the system.
    \end{itemize}
  \end{remark}
\end{frame}

\begin{frame}
  \begin{definition}[Transformation procedure, TP]
    \begin{itemize}
      \item Procedure which manipulates CDIs.
      \item Can take UDI as input, must convert to CDI.
    \end{itemize}
  \end{definition}

  \pause{}

  \begin{definition}[Integrity verification procedure, IVP]
    \begin{itemize}
      \item Checks the integrity of a CDI.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{block}{Certification rules}
    Should be checked so that the policy is consistent:
    \begin{description}
      \item[CR1] IVPs must ensure integrity of CDIs when IVPs are run.
      \item[CR2] TPs must be certified to be valid; valid CDIs transform into 
        valid CDIs; each TP can access restricted set of CDIs.
      \item[CR3] Access rules must satisfy separation-of-duties requirements.
      \item[CR4] All TPs must write to an append-only log.
      \item[CR5] Any TP handling UDI must convert it to a CDI or reject it.
    \end{description}
  \end{block}
\end{frame}

\begin{frame}
  \begin{block}{Enforcement rules}
    Describes the mechanisms needed in the system:
    \begin{description}
      \item[ER1] Must maintain and protect list of CDIs each TP can access.
      \item[ER2] Must maintain and protect list of TPs each subject can access.
      \item[ER3] The system must authenticate each subject requesting to execute 
        a TP\@.
      \item[ER4] Only a subject that may certify an access rule for a TP may 
        modify the respective entry in the list.
        This subject must not be allowed to execute this TP\@.
    \end{description}
  \end{block}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \small
  \printbibliography{}
\end{frame}

